---
title: "Aggregated Bus Dataset Under CSISG"
author: "Shermin Tan"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started

```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
packages = c('sp', 'rgdal', 'sf', 'tidyverse', 'tmap', 'ggplot2','stringr', 'maptools', 'raster', 'spatstat', 'sfheaders', 'dplyr')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```


#Import Data

```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
agg_bus <- read_csv("data/aspatial/agg_bus.csv")
```

#Data Cleaning and Segmentation

##Singapore Planning Area w/ Geometry Points

```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#transform CRS = 3414
mpsz_3414 <- st_set_crs(mpsz, 3414)

mpsz_3414_plnarea <- mpsz_3414 %>%
  group_by(PLN_AREA_N) %>%
  summarise()
```

###Visualize Maps Function
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Visualization
tmap_mode("view")
```

##agg_bus

###agg_bus overall for Year 2016 - 2019
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
agg_bus_overall <- agg_bus %>%
  dplyr::select(3,5:9)

agg_bus_overall_PA <- aggregate(agg_bus_overall, by = list(agg_bus_overall$planning_area), FUN=mean)
names(agg_bus_overall_PA)[1] <- 'planning_area'
agg_bus_overall_PA <- agg_bus_overall_PA %>%
  dplyr::select(1, 3:7)

agg_bus_overall_PA_mpsz3414 <- left_join(mpsz_3414_plnarea, agg_bus_overall_PA, by = c("PLN_AREA_N" = "planning_area"))
agg_bus_overall_PA_mpsz3414 <- agg_bus_overall_PA_mpsz3414 %>%
  group_by(PLN_AREA_N) %>%
  filter(! is.na(customer_expectations_score))
```

####Export to csv for PowerBI
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#agg_bus_overall_PA_mpsz3414_csv <- agg_bus_overall_PA_mpsz3414 %>%
#  dplyr::select(1,3:7) %>%
#  st_set_geometry(NULL)
#
#agg_bus_overall_PA_mpsz3414_csv$PLN_AREA_N <- sub("^", "Singapore, ", #agg_bus_overall_PA_mpsz3414_csv$PLN_AREA_N)
#
#write.csv(agg_bus_overall_PA_mpsz3414_csv, #"data/aspatial/agg_bus_overall_PowerBI.csv", row.names = FALSE)
```

####Visualization for agg_bus overall
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Customer Expectations Score
tm_shape(agg_bus_overall_PA_mpsz3414) + tm_fill("customer_expectations_score", style = "jenks", n= 6, title = "Average Customer Expectations Score (Bus) by Planning Area from 2016 - 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Satisfaction Score
tm_shape(agg_bus_overall_PA_mpsz3414) + tm_fill("customer_satisfaction_score", style = "jenks", n= 6, title = "Average Customer Satisfaction Score (Bus) by Planning Area from 2016 - 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Loyalty Trust Score
tm_shape(agg_bus_overall_PA_mpsz3414) + tm_fill("customer_loyalty‎_user_trust_score", style = "jenks", n= 6, title = "Average Customer Loyalty User Trust Score (Bus) by Planning Area from 2016 - 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Overall Quality Score
tm_shape(agg_bus_overall_PA_mpsz3414) + tm_fill("perceived_overall_quality_score", style = "jenks", n= 6, title = "Average Perceived Overall Quality Score by (Bus) Planning Area from 2016 - 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Value Score
tm_shape(agg_bus_overall_PA_mpsz3414) + tm_fill("perceived_value_score", style = "jenks", n= 6, title = "Average Perceived Value Score (Bus) by Planning Area from 2016 - 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```

####Expectations vs Satisfaction
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
agg_bus_overall_ExpectvsSatisf <- agg_bus_overall_PA_mpsz3414 %>%
  dplyr::select(1,2,3,6)

agg_bus_overall_ExpectvsSatisf_comparison <- agg_bus_overall_ExpectvsSatisf %>%
  mutate(PositiveOrNegative = case_when(customer_expectations_score > customer_satisfaction_score ~ "Yes", 
                                        customer_expectations_score < customer_satisfaction_score ~ "No",
                                        TRUE ~ "NA"))

agg_bus_overall_ExpectvsSatisf_positive <- agg_bus_overall_ExpectvsSatisf_comparison %>%
  filter(PositiveOrNegative == "Yes") %>%
  mutate(difference = customer_expectations_score - customer_satisfaction_score)

agg_bus_overall_ExpectvsSatisf_positive %>% arrange(desc(difference))

#Visualize difference
tm_shape(agg_bus_overall_ExpectvsSatisf_positive) + tm_fill("difference", style = "jenks", n= 6, title = "Bus:(AVG) Expectations vs Satisfaction Scores Difference (2016 - 2019)") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```

###Individually by year
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#2019
agg_bus_2019 <- agg_bus %>%
  filter(year == 2019) %>%
  dplyr::select(1,3,5:9)

agg_bus_2019_PA <- aggregate(agg_bus_2019, by = list(agg_bus_2019$planning_area), FUN=mean)
names(agg_bus_2019_PA)[1] <- 'planning_area'
agg_bus_2019_PA <- agg_bus_2019_PA %>%
  dplyr::select(1:2,4:8)

agg_bus_2019_PA_mpsz3414 <- left_join(mpsz_3414_plnarea, agg_bus_2019_PA, by = c("PLN_AREA_N" = "planning_area"))
agg_bus_2019_PA_mpsz3414 <- agg_bus_2019_PA_mpsz3414 %>%
  group_by(PLN_AREA_N) %>%
  filter(year == 2019)
  #drop_na()

#2018
agg_bus_2018 <- agg_bus %>%
  filter(year == 2018) %>%
  dplyr::select(1,3,5:9)

agg_bus_2018_PA <- aggregate(agg_bus_2018, by = list(agg_bus_2018$planning_area), FUN=mean)
names(agg_bus_2018_PA)[1] <- 'planning_area'
agg_bus_2018_PA <- agg_bus_2018_PA %>%
  dplyr::select(1:2,4:8)

agg_bus_2018_PA_mpsz3414 <- left_join(mpsz_3414_plnarea, agg_bus_2018_PA, by = c("PLN_AREA_N" = "planning_area"))
agg_bus_2018_PA_mpsz3414 <- agg_bus_2018_PA_mpsz3414 %>%
  group_by(PLN_AREA_N) %>%
  filter(year == 2018)
  #drop_na()

#2017
agg_bus_2017 <- agg_bus %>%
  filter(year == 2017) %>%
  dplyr::select(1,3,5:9)

agg_bus_2017_PA <- aggregate(agg_bus_2017, by = list(agg_bus_2017$planning_area), FUN=mean)
names(agg_bus_2017_PA)[1] <- 'planning_area'
agg_bus_2017_PA <- agg_bus_2017_PA %>%
  dplyr::select(1:2,4:8)

agg_bus_2017_PA_mpsz3414 <- left_join(mpsz_3414_plnarea, agg_bus_2017_PA, by = c("PLN_AREA_N" = "planning_area"))
agg_bus_2017_PA_mpsz3414 <- agg_bus_2017_PA_mpsz3414 %>%
  group_by(PLN_AREA_N)  %>%
  filter(year == 2017)
  #drop_na()

#2016
agg_bus_2016 <- agg_bus %>%
  filter(year == 2016) %>%
  dplyr::select(1,3,5:9)

agg_bus_2016_PA <- aggregate(agg_bus_2016, by = list(agg_bus_2016$planning_area), FUN=mean)
names(agg_bus_2016_PA)[1] <- 'planning_area'
agg_bus_2016_PA <- agg_bus_2016_PA %>%
  dplyr::select(1:2,4:8)

agg_bus_2016_PA_mpsz3414 <- left_join(mpsz_3414_plnarea, agg_bus_2016_PA, by = c("PLN_AREA_N" = "planning_area"))
agg_bus_2016_PA_mpsz3414 <- agg_bus_2016_PA_mpsz3414 %>%
  group_by(PLN_AREA_N)  %>%
  filter(year == 2016)
  #drop_na()
```

####Comparison of respective scores over the years
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
agg_bus_2016_2019_comparison <- agg_bus %>%
  dplyr::select(1,3,5:9)

agg_bus_2016_2019_comparison_PA <- aggregate(agg_bus_2016_2019_comparison, by = list(agg_bus_2016_2019_comparison$planning_area, agg_bus_2016_2019_comparison$year), FUN=mean)
names(agg_bus_2016_2019_comparison_PA)[1] <- 'planning_area'
names(agg_bus_2016_2019_comparison_PA)[2] <- 'year'
agg_bus_2016_2019_comparison_PA <- agg_bus_2016_2019_comparison_PA %>%
  dplyr::select(1:2,5:9)

agg_bus_2016_2019_comparison_PA <- agg_bus_2016_2019_comparison_PA %>%
  mutate_at("year", as.character)

#Customer Expectations Score
ggplot(agg_bus_2016_2019_comparison_PA, aes(planning_area, customer_expectations_score,  fill=year), xlab="Age Group") + geom_bar(stat="identity", position="dodge", width=.6) + scale_x_discrete(labels=abbreviate) + geom_text(aes(label=round(customer_expectations_score, 1)), vjust=0, size=3) + labs(title="Comparison of Customer Expectation Scores (Bus) from 2016 to 2019")

#Customer Satisfaction Score
ggplot(agg_bus_2016_2019_comparison_PA, aes(planning_area, customer_satisfaction_score,  fill=year), xlab="Age Group") + geom_bar(stat="identity", position="dodge", width=.6) + scale_x_discrete(labels=abbreviate) + geom_text(aes(label=round(customer_satisfaction_score, 1)), vjust=0, size=3) + labs(title="Comparison of Customer Satisfaction Scores (Bus) from 2016 to 2019")

#Customer Loyalty User Trust Score
ggplot(agg_bus_2016_2019_comparison_PA, aes(planning_area, agg_bus_2016_2019_comparison_PA$`customer_loyalty‎_user_trust_score`,  fill=year), xlab="Age Group") + geom_bar(stat="identity", position="dodge", width=.6) + scale_x_discrete(labels=abbreviate) + geom_text(aes(label=round(agg_bus_2016_2019_comparison_PA$`customer_loyalty‎_user_trust_score`, 1)), vjust=0, size=3) + labs(title="Comparison of Customer Loyalty User Trust Scores (Bus) from 2016 to 2019")

#Perceived Overall Quality Score
ggplot(agg_bus_2016_2019_comparison_PA, aes(planning_area, perceived_overall_quality_score,  fill=year), xlab="Age Group") + geom_bar(stat="identity", position="dodge", width=.6) + scale_x_discrete(labels=abbreviate) + geom_text(aes(label=round(perceived_overall_quality_score, 1)), vjust=0, size=3) + labs(title="Comparison of Perceived Overall Quality Scores (Bus) from 2016 to 2019")

#Perceived Value Score
ggplot(agg_bus_2016_2019_comparison_PA, aes(planning_area, perceived_value_score,  fill=year), xlab="Age Group") + geom_bar(stat="identity", position="dodge", width=.6) + scale_x_discrete(labels=abbreviate) + geom_text(aes(label=round(perceived_value_score, 1)), vjust=0, size=3) + labs(title="Comparison of Perceived Value Scores (Bus) from 2016 to 2019")
```

####Visualization for 2019 agg_bus
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Customer Expectations Score
tm_shape(agg_bus_2019_PA_mpsz3414) + tm_fill("customer_expectations_score", style = "jenks", n= 6, title = "Average Customer Expectations Score (Bus) by Planning Area in 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Satisfaction Score
tm_shape(agg_bus_2019_PA_mpsz3414) + tm_fill("customer_satisfaction_score", style = "jenks", n= 6, title = "Average Customer Satisfaction Score (Bus) by Planning Area in 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Loyalty Trust Score
tm_shape(agg_bus_2019_PA_mpsz3414) + tm_fill("customer_loyalty‎_user_trust_score", style = "jenks", n= 6, title = "Average Customer Loyalty User Trust Score (Bus) by Planning Area in 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Overall Quality Score
tm_shape(agg_bus_2019_PA_mpsz3414) + tm_fill("perceived_overall_quality_score", style = "jenks", n= 6, title = "Average Perceived Overall Quality Score (Bus) by Planning Area in 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Value Score
tm_shape(agg_bus_2019_PA_mpsz3414) + tm_fill("perceived_value_score", style = "jenks", n= 6, title = "Average Perceived Value Score (Bus) by Planning Area in 2019") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```

####Visualization for 2018 agg_bus
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Customer Expectations Score
tm_shape(agg_bus_2018_PA_mpsz3414) + tm_fill("customer_expectations_score", style = "jenks", n= 6, title = "Average Customer Expectations Score (Bus) by Planning Area in 2018") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Satisfaction Score
tm_shape(agg_bus_2018_PA_mpsz3414) + tm_fill("customer_satisfaction_score", style = "jenks", n= 6, title = "Average Customer Satisfaction Score (Bus) by Planning Area in 2018") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Loyalty Trust Score
tm_shape(agg_bus_2018_PA_mpsz3414) + tm_fill("customer_loyalty‎_user_trust_score", style = "jenks", n= 6, title = "Average Customer Loyalty User Trust Score (Bus) by Planning Area in 2018") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Overall Quality Score
tm_shape(agg_bus_2018_PA_mpsz3414) + tm_fill("perceived_overall_quality_score", style = "jenks", n= 6, title = "Average Perceived Overall Quality Score (Bus) by Planning Area in 2018") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Value Score
tm_shape(agg_bus_2018_PA_mpsz3414) + tm_fill("perceived_value_score", style = "jenks", n= 6, title = "Average Perceived Value Score (Bus) by Planning Area in 2018") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```

####Visualization for 2017 agg_bus
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Customer Expectations Score
tm_shape(agg_bus_2017_PA_mpsz3414) + tm_fill("customer_expectations_score", style = "jenks", n= 6, title = "Average Customer Expectations Score (Bus) by Planning Area in 2017") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Satisfaction Score
tm_shape(agg_bus_2017_PA_mpsz3414) + tm_fill("customer_satisfaction_score", style = "jenks", n= 6, title = "Average Customer Satisfaction Score (Bus) by Planning Area in 2017") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Loyalty Trust Score
tm_shape(agg_bus_2017_PA_mpsz3414) + tm_fill("customer_loyalty‎_user_trust_score", style = "jenks", n= 6, title = "Average Customer Loyalty User Trust Score (Bus) by Planning Area in 2017") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Overall Quality Score
tm_shape(agg_bus_2017_PA_mpsz3414) + tm_fill("perceived_overall_quality_score", style = "jenks", n= 6, title = "Average Perceived Overall Quality Score (Bus) by Planning Area in 2017") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Value Score
tm_shape(agg_bus_2017_PA_mpsz3414) + tm_fill("perceived_value_score", style = "jenks", n= 6, title = "Average Perceived Value Score (Bus) by Planning Area in 2017") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```

####Visualization for 2016 agg_bus
```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#Customer Expectations Score
tm_shape(agg_bus_2016_PA_mpsz3414) + tm_fill("customer_expectations_score", style = "jenks", n= 6, title = "Average Customer Expectations Score (Bus) by Planning Area in 2016") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Satisfaction Score
tm_shape(agg_bus_2016_PA_mpsz3414) + tm_fill("customer_satisfaction_score", style = "jenks", n= 6, title = "Average Customer Satisfaction Score (Bus) by Planning Area in 2016") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Customer Loyalty Trust Score
tm_shape(agg_bus_2016_PA_mpsz3414) + tm_fill("customer_loyalty‎_user_trust_score", style = "jenks", n= 6, title = "Average Customer Loyalty User Trust Score (Bus) by Planning Area in 2016") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Overall Quality Score
tm_shape(agg_bus_2016_PA_mpsz3414) + tm_fill("perceived_overall_quality_score", style = "jenks", n= 6, title = "Average Perceived Overall Quality Score (Bus) by Planning Area in 2016") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)

#Perceived Value Score
tm_shape(agg_bus_2016_PA_mpsz3414) + tm_fill("perceived_value_score", style = "jenks", n= 6, title = "Average Perceived Value Score (Bus) by Planning Area in 2016") + tm_text("PLN_AREA_N", size = 0.65) + tm_shape(mpsz_3414) + tm_borders(lwd = 0.1, alpha = 1) + tm_legend(legend.outside=TRUE) #+ tm_text("PLN_AREA_N", size = 0.2)
```
































